---
title: "Comparison of pop-gen stats across tools"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library("here")
library("ggplot2")
library("cowplot")
library("reshape")
library("plyr")
library("ggthemes")
library("ggpol")
```

# 1. Read in data.
```{r}
# Read in full samples data
pops_data<-read.csv(here::here("rdata","dest_v2.samps_8Jun2023.csv"),header=TRUE)
pops<-pops_data$sampleId
# Keep only non-SIM populations
pops<-pops[grep("SIM",pops,invert = TRUE)]

# Read in the genome summary data.
genome_data<-read.csv(here::here("rdata","chroms_file.csv"),header=FALSE)
colnames(genome_data)<-c("chrom","len")

win<-"win10000" # Change this depending on which window size to read in (win10000, win50000, or win100000).

# Read in EggLib results
load(here::here("rdata/egglib",paste("egglib_",win,".RData",sep="")))
egglib_pop_genome_stats<-pop_genome_stats
egglib_pop_genome_stats$pi_sites<-egglib_pop_genome_stats$pi/egglib_pop_genome_stats$n_goodsites
head(egglib_pop_genome_stats,n=50)

# Read in npstat results
load(here::here("rdata/npstat",paste("npstat_",win,".RData",sep="")))
npstat_pop_genome_stats<-pop_genome_stats
npstat_pop_genome_stats$pi_sites<-npstat_pop_genome_stats$pi/npstat_pop_genome_stats$chrom_good_len
head(npstat_pop_genome_stats,n=50)
```

# 2. Compare estimates of pi across estimators (EggLib vs. npstat)
## 2.1 Summarise pi across the genome.
Exclude X chromosome.
```{r}
# EggLib
egglib_pop_genome_stats_summary<-ddply(egglib_pop_genome_stats[which(egglib_pop_genome_stats$chrom!="X"),],
                                .variables = c("pop"),summarise,
                                "pi_wt_mean"=weighted.mean(pi_sites,chrom_len/sum(chrom_len)),
                                "tajd_wt_mean"=weighted.mean(med_tajD,chrom_len/sum(chrom_len)),
                                "n_good_sites"=sum(n_goodsites))
# Add continent info
egglib_pop_genome_stats_summary$continent<-pops_data$continent[
  match(egglib_pop_genome_stats_summary$pop,pops_data$sampleId)]
head(egglib_pop_genome_stats_summary)

# npstat
npstat_pop_genome_stats_summary<-ddply(npstat_pop_genome_stats[which(npstat_pop_genome_stats$chrom!="X"),],
                                .variables = c("pop"),summarise,
                                "pi_wt_mean"=weighted.mean(pi_sites,chrom_len/sum(chrom_len)),
                                "tajd_wt_mean"=weighted.mean(med_tajD,chrom_len/sum(chrom_len)),
                                "n_good_sites"=sum(chrom_good_len))
# Add continent info
npstat_pop_genome_stats_summary$continent<-pops_data$continent[
  match(npstat_pop_genome_stats_summary$pop,pops_data$sampleId)]
head(npstat_pop_genome_stats_summary)
```

## 2.1 Combine data
```{r}
pop_genome_stats_pi<-data.frame("ID"=pops_data$sampleId[which(pops_data$sampleId!="SIM_SIM_w501_1_NA-MM-DD")],
                                "Continent"=pops_data$continent[which(pops_data$sampleId!="SIM_SIM_w501_1_NA-MM-DD")])

pop_genome_stats_pi$pi_egglib<-egglib_pop_genome_stats_summary$pi_wt_mean[
  match(pop_genome_stats_pi$ID,egglib_pop_genome_stats_summary$pop)]
pop_genome_stats_pi$tajD_egglib<-egglib_pop_genome_stats_summary$tajd_wt_mean[
  match(pop_genome_stats_pi$ID,egglib_pop_genome_stats_summary$pop)]
pop_genome_stats_pi$n_goodsites_egglib<-egglib_pop_genome_stats_summary$n_good_sites[
  match(pop_genome_stats_pi$ID,egglib_pop_genome_stats_summary$pop)]

pop_genome_stats_pi$pi_npstat<-npstat_pop_genome_stats_summary$pi_wt_mean[
  match(pop_genome_stats_pi$ID,npstat_pop_genome_stats_summary$pop)]
pop_genome_stats_pi$tajD_npstat<-npstat_pop_genome_stats_summary$tajd_wt_mean[
  match(pop_genome_stats_pi$ID,npstat_pop_genome_stats_summary$pop)]
pop_genome_stats_pi$n_goodsites_npstat<-npstat_pop_genome_stats_summary$n_good_sites[
  match(pop_genome_stats_pi$ID,npstat_pop_genome_stats_summary$pop)]

head(pop_genome_stats_pi)
```

## 2.2 Comparison
Remove samples for which not all tools give estimates
```{r}
pop_genome_stats_pi<-pop_genome_stats_pi[
  which(!(pop_genome_stats_pi$pi_egglib==0 | pop_genome_stats_pi$pi_npstat==0)),]
pop_genome_stats_pi<-pop_genome_stats_pi[complete.cases(pop_genome_stats_pi),]
```

Add some statistics from metadata file
```{r}
pop_genome_stats_pi$Cov<-pops_data$Cov[match(pop_genome_stats_pi$ID,pops_data$sampleId)]
```

How many samples have estimates in all tools
```{r}
head(pop_genome_stats_pi)
nrow(pop_genome_stats_pi)
```

Figure 1:
Compare PI across estimators
```{r}
pop_genome_stats_pi[pop_genome_stats_pi$Continent=="North_America",]$Continent <- "North America"
pop_genome_stats_pi[pop_genome_stats_pi$Continent=="South_America",]$Continent <- "South America"

npstat_v_egglib <- ggplot() +
    geom_point(data = pop_genome_stats_pi, aes(x = pi_npstat, y = pi_egglib, colour = Continent), size = 3) +
    xlab(expression(paste(pi, " npstat", sep = ""))) +
    ylab(expression(paste(pi, " EggLib", sep = ""))) +
    theme(text = element_text(size = 16)) + 
    scale_color_gdocs() + 
    theme_bw() + 
    geom_smooth(data = pop_genome_stats_pi, aes(x = pi_npstat, y = pi_egglib, group = 0), method = "lm", color = "black", se = F) +
    annotate("text", x = 0.003, y = Inf, label = "Spearman's Ï = 0.87, P < 0.001\nPearson's cor. = 0.82, P < 0.001", hjust = 0, vjust = 1) 
npstat_v_egglib
```

```{r}
cor.test(x=pop_genome_stats_pi$pi_npstat,
         y=pop_genome_stats_pi$pi_egglib,method="spearman")

cor.test(x=pop_genome_stats_pi$pi_npstat,
         y=pop_genome_stats_pi$pi_egglib,method="pearson")
```

Figure 2:
Plot pi by continent for all estimators
```{r}
pop_genome_stats_pi_m<-melt(pop_genome_stats_pi,measure.vars = c("pi_egglib","pi_npstat"))
pop_genome_stats_pi_m$variable<-factor(pop_genome_stats_pi_m$variable,levels=c("pi_egglib","pi_npstat"),labels=c("EggLib 3","npStat"))
head(pop_genome_stats_pi_m)

partb<-ggplot(data=pop_genome_stats_pi_m,aes(x=Continent,y=value,color=Continent))+
    #geom_boxplot(data=pop_genome_stats_pi_m,aes(x=Continent,y=value,fill=Continent), show.legend = F)+
    geom_boxjitter(outlier.colour = NA, outlier.fill = NA, alpha = 0.5, show.legend = F)+
    ylab(expression(paste(pi)))+
    facet_grid(.~variable,scales="free_y")+
    theme_bw() +
    theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1),
          text=element_text(size=16,color="black"),axis.text = element_text(color="black")) + labs(x="")  + 
    scale_fill_gdocs() + scale_color_gdocs()
partb
cowplot::plot_grid(npstat_v_egglib,partb,ncol = 1, align = "l",labels="AUTO")
ggsave("FigureS5.pdf",plotS5,height = 8, width = 6)
```




_______________________________________________________________________________
_______________________________________________________________________________
Below are some additional comparisons. 
Not included in the figures in the manuscript.

# 3. Compare n_good_sites across tools
```{r}
npstat_v_egglib<-ggplot()+
  geom_point(data=pop_genome_stats_pi[which(pop_genome_stats_pi$Cov > 20 & pop_genome_stats_pi$Cov < 100),],
             aes(x=n_goodsites_npstat/1000000,y=n_goodsites_egglib/1000000,colour=Continent),size=3)+
  xlab("N good sites npstat")+
  ylab("N good sites EggLib")+
  theme(text=element_text(size=16))+
  facet_wrap(Continent~.,scales="free")
npstat_v_egglib
```

```{r}
cor.test(x=pop_genome_stats_pi$n_goodsites_npstat,
         y=pop_genome_stats_pi$n_goodsites_egglib,method="spearman")
cor.test(x=pop_genome_stats_pi$n_goodsites_npstat,
         y=pop_genome_stats_pi$n_goodsites_egglib,method="pearson")
```

Low agreement overall, why is this?
```{r}
pop_genome_stats_pi[
  which(pop_genome_stats_pi$Continent=="North_America" & 
          pop_genome_stats_pi$n_goodsites_npstat < 90000000 & 
          pop_genome_stats_pi$Cov > 20 & 
          pop_genome_stats_pi$Cov < 100),]
```

Is the difference in n good sites related to coverage?
```{r}
# If negative, egglib has more good sites
pop_genome_stats_pi$n_sites_diff<-pop_genome_stats_pi$n_goodsites_npstat-pop_genome_stats_pi$n_goodsites_egglib
ggplot()+
  geom_point(data=pop_genome_stats_pi,
             aes(x=Cov,y=n_sites_diff,colour=Continent),size=3)+
  xlab("Cov")+
  ylab("Difference in N good sites (npstat-EggLib)")+
  theme(text=element_text(size=16))+
  facet_wrap(Continent~.)
```


Is this also explaining the off-diagonal cloud of points in the pi comparison?
```{r}
# If negative, egglib has higher pi
pop_genome_stats_pi$pi_diff<-pop_genome_stats_pi$pi_npstat-pop_genome_stats_pi$pi_egglib
ggplot()+
  geom_point(data=pop_genome_stats_pi,
             aes(x=Cov,y=pi_diff,colour=Continent),size=3)+
  xlab("Cov")+
  ylab(expression(paste("Difference in",pi," (npstat-EggLib)")))+
  theme(text=element_text(size=16))+
  facet_wrap(Continent~.)

npstat_v_egglib<-ggplot()+
  geom_point(data=pop_genome_stats_pi,aes(x=pi_npstat,y=pi_egglib,colour=Continent),size=3)+
  geom_point(data=pop_genome_stats_pi[which(pop_genome_stats_pi$Cov < 20 | pop_genome_stats_pi$Cov > 100),],
             aes(x=pi_npstat,y=pi_egglib),size=3,shape=21,colour="red",stroke=2)+
  xlab(expression(paste(pi," npstat",sep="")))+
  ylab(expression(paste(pi," EggLib",sep="")))+
  theme(text=element_text(size=16))
npstat_v_egglib
```

# 4. Tajima's D Comparison

Plot Tajima's D by continent for all estimators
```{r}
pop_genome_stats_tajD_m<-melt(pop_genome_stats_pi,measure.vars = c("tajD_egglib","tajD_npstat"))
pop_genome_stats_tajD_m$variable<-factor(pop_genome_stats_tajD_m$variable,
                                         levels=c("tajD_egglib","tajD_npstat"),labels=c("EggLib","npstat"))
head(pop_genome_stats_tajD_m)
ggplot()+
  geom_boxplot(data=pop_genome_stats_tajD_m,aes(x=Continent,y=value,fill=Continent))+
  geom_point(data=pop_genome_stats_tajD_m,aes(x=Continent,y=value),position=position_jitter(0.1))+
  ylab(expression(paste(pi)))+
  facet_wrap(.~variable,scales="free_y")+
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5,size=16),
        text=element_text(size=16))
```

How many samples have estimates in all tools
```{r}
head(pop_genome_stats_pi)
nrow(pop_genome_stats_pi)
```

Compare Tajima's D across estimators
```{r}
npstat_v_egglib<-ggplot()+
  geom_point(data=pop_genome_stats_pi,aes(x=tajD_npstat,y=tajD_egglib,colour=Continent),size=3)+
  xlab("Tajima's D npstat")+
  ylab("Tajima's D EggLib")+
  theme(text=element_text(size=16))
npstat_v_egglib
```

```{r}
cor.test(x=pop_genome_stats_pi$tajD_npstat,
         y=pop_genome_stats_pi$tajD_egglib,method="spearman")
cor.test(x=pop_genome_stats_pi$tajD_npstat,
         y=pop_genome_stats_pi$tajD_egglib,method="pearson")
```

