---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(UpSetR)
library(ComplexUpset)
library(tidyverse)
library(reshape2)
```

Modified fromList: [https://github.com/hms-dbmi/UpSetR/issues/85#issuecomment-327900647](https://github.com/hms-dbmi/UpSetR/issues/85#issuecomment-327900647)
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```

# 1. Data processing
## 1.1 Gene sets from DEST v. 2
```{r}
xtx_genes<-scan(here::here("rdata/dest2_seasonality","xtx_outlier_gowinda_genes.txt"),
                what="character")
xtx_genes<-gsub("fbgn","FBgn",xtx_genes)
c2_genes<-scan(here::here("rdata/dest2_seasonality","c2_outlier_gowinda_genes.txt"),
                what="character")
c2_genes<-gsub("fbgn","FBgn",c2_genes)
# All genes in windows with Fgt > 0.2 in any pairwise comparison
fgt_genes<-scan(here::here("rdata/dest2_seasonality","fgt_outlier_gowinda_genes.txt"),
                what="character")
fgt_genes<-unique(fgt_genes)
fgt_genes<-gsub("fbgn","FBgn",fgt_genes)
```

## 1.2 Gene sets from other studies
Compare set of genes in outlier regions to genes identified in other publications:

1. Bogaerts Márqez et al. 2020 Mol. Ecol. 30: 938-954
  Genes associated with environmental variables

2. Machado et al. 2015 Mol. Ecol. 25: 723-740
  Genes clinal in D. mel and D. sim

3. Rudman et al. 2022 Science. 375: eabj7484.
  Genes near outlier SNPs showing rapid temporal alelle frequency variation

4. Rajpurohit et al. 2018. Mol. Ecol. 27: 3525-3540.
  Genes with SNPs associated with dessication resistance in 175 DGRP lines.
  
5. Erickson et al. 2020
  Genes near top 0.1% of SNPs associated with two measures of diapause.
  [N.B. many (!) malformatted gene IDs that cannot be converted to FBgn IDs]
  
6. Machado et al. 2021
  Genes overlapping with top 1% of seasonal SNPs
  
7. Nunez et al. 2023
  Genes overlapping with SNPs with rank normalised p-values < 0.05. 

```{r}
# 1
bogaerts_marquez_et_al_2020<-read.table(
  here::here("rdata/dest2_seasonality/Bogaerts_Márquez_et_al_2021/","tableS10.csv"),
  header=TRUE,sep=",")
# 2
machado_2015<-read.table(here::here("rdata/dest2_seasonality/Machado_et_al_2015",
                                    "sharedgenes_FlyBase_Fields_download.txt"),
                         header=FALSE)
length(unique(machado_2015$V1))
length(unique(machado_2015$V2))
machado_2015<-unique(machado_2015$V2)

# 3
rudman_2022<-scan(here::here("rdata/dest2_seasonality/Rudman_et_al_2022","outlier_genes.txt"),
                  what="character")
# 4
rajpurohit_2018<-scan(here::here("rdata/dest2_seasonality/Rajpurohit_et_al_2018","topsnps_genes.txt"),
                      what="character")
rajpurohit_2018<-unique(rajpurohit_2018)

# 5
erickson_2020_0.1<-read.table(
  here::here("rdata/dest2_seasonality/Erickson_et_al_2020",
             "uniq_0.1perc_genes_FlyBase_Fields_download.txt"),
  fill = TRUE,quote = "",header=TRUE,sep="\t")

# 6
machado_2021<-scan(here::here("rdata/dest2_seasonality/Machado_et_al_2021",
                              "machado_2021_seasonalsnps_genes.txt"),
                   what="character")
machado_2021<-unique(machado_2021)

# 7
final.windows.pos = 
  data.frame(win.name = c("win_3.1", "win_4.7", "win_5.1", "win_6.1", "win_6.8", "win_9.6"),
             mid = c(3.0, 4.67, 5.12, 6.2, 6.8 , 9.6),
             chr = "2L") %>%
  mutate(start = (mid-0.2)*1e6 ,
         end  = (mid+0.2)*1e6  )

write.table(final.windows.pos[,c("chr","start","end")],
            here::here("rdata/dest2_seasonality/Nunez_et_al_2024","outlier_windows.bed"),
            quote=FALSE,col.names=FALSE,row.names=FALSE,sep="\t")

nunez_2024<-scan(here::here("rdata/dest2_seasonality/Nunez_et_al_2024",
                            "nunes_2024_outlier_windows_genes.txt"),
                   what="character")
nunez_2024<-unique(nunez_2024)
```

#### Process and unify the results
```{r}
l<-list("Dest_v2_xtx"=xtx_genes,
     "Dest_v2_c2"=c2_genes,
#     "Dest_v2_fgt"=fgt_genes,
     "Rudman_et_al_2022"=rudman_2022,
     "Machado_et_al_2015"=machado_2015,
     "Rajpurohit_et_al_2018"=rajpurohit_2018,
     "Erickson_et_al_2020"=unique(erickson_2020_0.1$FBID_KEY),
     "Bogaerts_Márquez_et_al_2021"=bogaerts_marquez_et_al_2020$FlyBase.code,     
     "Machado_et_al_2021"=machado_2021,
     "Nunez_et_al_2024"=nunez_2024)
lapply(l,FUN = length)
d<-fromList(l)

d$id<-as.character(1:nrow(d))
```

Save as Rdata object
```{r}
save(d,l,file=here::here("rdata/dest2_seasonality/","overlaps.RData"))
```

# 2. Overlaps analyis -- including Bergland 2016
#### NEW
#### 8. Extracting data from Bergland et al.
```{r}
load("/netfiles/nunezlab/D_melanogaster_resources/Datasets/2016.Bergland/6d_data.Rdata")

cbind(p, info) -> 

p_info %>%
filter(sfsfsfX.q<.3) ->
  seasonal_hits

c(
seasonal_hits$INTRON_SPLICE, 
seasonal_hits$STOP,
seasonal_hits$UTR_3_PRIME, 
seasonal_hits$UTR_5_PRIME, 
seasonal_hits$START,  
seasonal_hits$NS,
seasonal_hits$S
) %>% .[complete.cases(.)] -> gene_id_hits

unique(gene_id_hits) -> genes_of_interest_bergland_2016

write.table(genes_of_interest_bergland_2016, 
            file = "genes_of_interest_bergland_2016.txt", 
            append = FALSE, quote = FALSE, sep = "\t",
            eol = "\n", na = "NA", dec = ".", row.names = FALSE,
            col.names = FALSE, qmethod = c("escape", "double"),
            fileEncoding = "")

B2016_genes = read.delim2("/netfiles/nunezlab/D_melanogaster_resources/Datasets/2016.Bergland/Bergland2016_outliers_FBGNs.txt")

B2016_genes$FBID_KEY[-which(B2016_genes$FBID_KEY == "")] -> B2016_genes_fbgns
```
#### NEW ^^

Load data
```{r}
load("overlaps.RData")
l$Bergland_et_al_2016 <- B2016_genes_fbgns

d<-fromList(l)

d$id<-as.character(1:nrow(d))

```


```{r}
# Only those in at least 2 sets
## This will remove those that are unique to other sets
# d[rowSums(d[names(l)]) > 1,]

# Only those in Dest_v2_c2 OR Dest_v2_xtx OR Dest_v2_fgt AND at least 1 other set
#d[which(rowSums(d[names(l)[1:3]]) >= 1 & rowSums(d[names(l)[3:length(l)]]) >= 1),]

# Only those that are in either Dest_v2_c2 OR Dest_v2_xtx OR Dest_v2_fgt
## This will add those that are unique to Dest_v2_c2 OR Dest_v2_xtx OR Dest_v2_fgt
## this will still collect some that also occur in other sets so we need to remove duplicates after
# d[rowSums(d[names(l)[1:3]]) == 1,]

d_reduced<-rbind(d[which(rowSums(d[names(l)[1:2]]) >= 1 & rowSums(d[names(l)[3:length(l)]]) >= 1),],
                 d[rowSums(d[names(l)[1:2]]) >= 1,])

# need to remove duplicates
d_reduced<-d_reduced[which(!duplicated(d_reduced$id)),names(l)]

# Save this table
write.table("overlaps_genes_table.tab",
            col.names=TRUE,row.names=TRUE,quote=FALSE,sep="\t")


sum(d_reduced$Dest_v2_xtx)
sum(d_reduced$Dest_v2_c2)

names(d_reduced)[which(names(d_reduced) %in% c("Dest_v2_xtx","Dest_v2_c2") )] =
  c("z.Dest_v2_xtx", "z.Dest_v2_c2")

names(l)[which(names(l) %in% c("Dest_v2_xtx","Dest_v2_c2") )] =
  c("z.Dest_v2_xtx", "z.Dest_v2_c2")
```

Main plot
```{r}
ComplexUpset::upset(d_reduced,intersect = names(l),sort_sets=FALSE,
      base_annotations=list(
        'Intersection size'=intersection_size(
            text=list(
                vjust=0.5,
                hjust=-0.1,
                angle=90)
            )),width_ratio = 0.1,set_sizes = FALSE,keep_empty_groups=TRUE) ->
  upset_all
ggsave(upset_all, file ="upset_all.pdf")

```

Same plot but wih a min intersection size of 10
```{r}
ComplexUpset::upset(d_reduced,intersect = sort(names(l), decreasing = TRUE),
                    min_size=10,
                    stripes=c('cornsilk1', 'deepskyblue1', 'grey90'),
                    sort_sets=FALSE,
      base_annotations=list(
        'Intersection size'=intersection_size(
            text=list(
                vjust=0.5,
                hjust=-0.1,
                angle=90)
            )),width_ratio = 0.1,set_sizes = FALSE)->
  upset_10
ggsave(upset_10, file ="upset_all.pdf", h = 4, w = 6)

```

